---
title: "Boilerplate detection: template links that live in the content"
author: "Bart Turczynski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Boilerplate detection: template links that live in the content}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pagerankr)
```

## The links placement cannot reach

Weighting links by page region — `vignette("presets")`, `preset = "content"` —
handles site chrome. Navigation, header, footer and sidebar stop voting at full
strength, and the ranking stops being manufactured by whatever the template
repeats on every page.

It handles chrome and nothing else, because it can only ask *where on the page
is this link*. A recycled call-to-action inside the article body, a compliance
link dropped into every post, an author byline under every headline — these are
in the content region. They *are* content, structurally. Placement will keep
them at full weight forever, and they carry exactly the problem chrome does:
one editorial decision, made once in a template, cast as thousands of votes.

The boilerplate detector is the second way of asking. Instead of *where is this
link*, it asks **does this component always point at the same place**.

## The metric

The unit is a **container**: the template element a link sits in, identified
stably across the pages it appears on. For each container-and-target pair:

- **denominator** — the number of pages the container appears on at all;
- **numerator** — the number of those pages where it points at *this* target;
- **ratio** — numerator over denominator.

`ratio` is a boilerplate score in `[0, 1]`, and **higher means more
boilerplate**. The polarity is easy to invert when reading, so it is worth
stating twice:

- `ratio = 1.0` — every single time this component appeared, it linked here. A
  template link. Gets discounted.
- `ratio → 0` — this component picks a different target on each page. A genuine
  editorial choice, made per page. Keeps full weight.

That separates two things which look identical structurally:

| Component | Behavior | Ratio | Verdict |
|---|---|---|---|
| A recycled "Book a demo" CTA | always the same target | ~1.0 | boilerplate |
| A related-posts module | different articles per page | low | **not** boilerplate |

Both recur identically across the site. Only the first one is a template
casting the same vote over and over; the second is the template asking a real
question and getting a different answer every time. A detector keyed on "this
component repeats" would flag both, and would be wrong about the second.

### Why the container is the denominator

Scoring against the whole site instead would flag the homepage — correctly for
the header logo, wrongly for an in-body link from an article that genuinely
chose to point there. Same destination, opposite nature. **Boilerplate is a
property of the edge and of its source context, not of the destination.**

Conditioning on the container also keeps the metric local. Landing pages, blog
categories and post templates each have their own component sets, and a ratio
computed within one of them says something a sitewide count cannot.

## Turning it on

The detector is off by default. It is switched on by **data** — the
`container_col` argument naming a column of container identities — in the same
way `placement_col` switches on region weighting and `nofollow_col` switches on
nofollow handling.

Here is a twelve-post blog. Every post carries a byline component linking one
author, and a related-posts component linking a different post each time:

```{r}
posts <- sprintf("/post-%02d", 1:12)

edges <- rbind(
  # The byline: same component, same target, every single page.
  data.frame(from = posts, to = "/author/dana", container = "byline"),
  # Related posts: same component, a different target on each page.
  data.frame(from = posts, to = rev(posts), container = "related"),
  # A CTA that mostly, but not always, points at pricing.
  data.frame(
    from = posts,
    to = c(rep("/pricing", 7), sprintf("/guide-%02d", 1:5)),
    container = "cta"
  )
)

scored <- pagerank(edges, container_col = "container")
head(scored[order(-scored$pagerank), ], 4)
```

All three components appear on all twelve pages, and the ratio tells them
apart. The byline scores `12/12 = 1.0` and is discounted. `related` scores
`1/12` per target and keeps full weight. The CTA scores `7/12 = 0.58` for
`/pricing` — over the default threshold, so discounted — and `1/12` for each
guide, which is not.

The provenance is recorded in the transition audit, so a run can always account
for what it discounted and why:

```{r}
attr(scored, "transition_audit")$config$boilerplate
```

### It is a downweight, never a drop

Discounted edges stay in the graph at reduced weight. This is the same rule
that governs region weighting, for the same two reasons: dropping edges changes
the graph's *shape* rather than its transition probabilities — pages reachable
only through a template become teleport-only, pages linking out only through
one become dangling — and detection is a heuristic, so a misclassified link at
half weight is a small error where a deleted one is silent.

## The four arguments

| Argument | Default | What it does |
|---|---|---|
| `container_col` | `NULL` | Names the container column. **Supplying it turns the detector on.** |
| `boilerplate_threshold` | `0.5` | Ratio at or above which an edge is *classified* boilerplate. |
| `min_container_pages` | `10` | Pages a container must appear on before any of its edges may be classified at all. |
| `boilerplate_weight` | `0.5` | The multiplier applied *once* an edge is classified. |

**`boilerplate_threshold` and `boilerplate_weight` are unrelated quantities
that happen to share a default of 0.5.** One is a fraction of pages that
decides *whether* an edge is boilerplate; the other is the discount applied
once it is. Never read a bare "0.5" in this area without checking which one it
refers to.

`0.5` as a threshold is a documented default, not an empirical cut — there is
no natural break in the ratio distribution to find. It sits where it does
because a stricter `0.9` misses two whole families of real boilerplate that
recur across every crawl tested: recurring in-content CTAs (ratios 0.54–0.82)
and author byline links (0.53–0.69).

`min_container_pages` is an evidence floor. A container appearing on three
pages can only score `0.33`, `0.67` or `1.0`, so a high ratio there is
quantization rather than signal. Ten is a judgement call, not a measured cut —
raise it if your containers are large, lower it if your site is small.

The CTA above is exactly the case the threshold decides. At the default it is
caught; at `0.9` it is not, and only the byline is:

```{r}
discounted <- function(threshold) {
  run <- pagerank(edges, container_col = "container",
                  boilerplate_threshold = threshold)
  attr(run, "transition_audit")$config$boilerplate$n_edges_discounted
}

c(default = discounted(0.5), strict = discounted(0.9))
```

Seven of those nineteen edges are the recurring CTA. On a real site that is the
difference between catching the promotional module on every post and missing
it — which is why the default sits at `0.5` rather than somewhere stricter and
more comfortable-looking.

## Composing with region weighting

Region and recurrence are **two detectors feeding one axis**, not two axes.

This matters arithmetically. A nav link is boilerplate *by construction* — it
is the same fact detected twice — so multiplying the two factors would discount
it to `0.1 × 0.5 = 0.05`, a number nobody can explain and which no one
intended. Instead the **strongest applicable discount wins**:

| Edge | Detected by | Weight |
|---|---|---:|
| nav / header / footer | region | 0.10 |
| repetitive in-content | recurrence | 0.50 |
| unique in-content | neither | 1.00 |

Three tiers, and chrome does not get pushed into a fourth just because two
detectors agree about it. Supply both `placement_col` and `container_col` and
you get exactly this table:

```{r}
mixed <- data.frame(
  from = rep(posts, each = 2),
  to = rep(c("/home", "/author/dana"), times = 12),
  region = rep(c("nav", "content"), times = 12),
  container = rep(c("mainnav", "byline"), times = 12)
)

both <- pagerank(
  mixed,
  placement_col = "region",
  placement_weights = c(content = 1, nav = 0.1, header = 0.1,
                        footer = 0.1, aside = 0.1),
  container_col = "container"
)
head(both[order(-both$pagerank), ], 3)
```

The nav edge stays at `0.1` rather than falling to `0.05`, and the byline —
which region weighting could never have touched — lands at `0.5`.

## What to expect on a real site: author pages will drop

This is the consequence worth knowing in advance, because it looks like a bug
the first time you see it.

Byline links are textbook boilerplate: one template decision, replicated across
every article, pointing at the same author page each time. On one crawl a
single author page was linked from 4,116 of 7,563 pages by an identical
element. Under an unweighted PageRank that in-degree makes author pages some of
the strongest nodes on the site.

Turn the detector on and they fall — often a long way.

**That is the detector working, not failing.** None of those thousands of links
was an editorial judgment that the author page deserved authority; one of them
was, and it got replicated by a template. The ranking those links produced was
manufactured by the markup rather than earned by the page. Discounting them
puts author pages roughly where their genuinely editorial inbound links place
them.

The same reasoning applies to the other families the detector reliably finds —
terms and privacy pages, cart and account links, promotional CTAs. If a page's
rank drops sharply when you enable the detector, that is a measurement of how
much of its rank came from a template.

If you disagree in a specific case, the judgement is yours to make: raise
`boilerplate_threshold`, raise `boilerplate_weight` toward 1, or curate the
container column so the component in question is not scored. The detector is a
convenience, not an oracle, and "boilerplate but in main content" is ultimately
a call about your own site.

## Getting a container column from a crawl

`container_col` is crawler-neutral by design: `pagerank()` only consumes
container *identity* and never asks how it was derived. Any crawler that can
say "these two links are the same template element" can drive the detector — a
DOM path, a CSS selector, a template ID from a CMS export.

For Screaming Frog, `sf_container_from_path()` derives identity from the DOM
path SF exports:

```{r}
sf_container_from_path(c(
  "//body/div/main/article/div[@class='byline']/a[1]",
  "//body/div/main/article/div[@class='byline']/a[3]",
  "//body/div/main/article/p[5]/a"
))
```

Two things happen there. The link's own step is dropped, so the first and third
anchor inside one byline resolve to the **same** container rather than to two —
which is the point, since they are one component. And numeric indices are
stripped, because the same component lands at `p[5]` on a long post and `p[3]`
on a short one, so positions are noise for this question. Class predicates
survive, because a class is exactly the stable component identifier we want.

On real crawls this compresses 22,022 raw paths to 1,630 skeletons. The
normalization is load-bearing: without it the detector under-detects in-content
components while working fine on navigation, which is precisely backwards —
navigation is already covered by region weighting.

Note this cuts the **opposite** way from `sf_region_from_path()`, which strips
predicates entirely so that a `div[@class='site-footer']` is not mistaken for a
`<footer>` element. The two functions answer different questions — *which
region is this* versus *is this the same component* — and the difference is
deliberate.

`screaming_frog_links()` carries a `container` column for you. Note that
`pagerank_screaming_frog()` does **not** pass it automatically: enabling a
detector that reshuffles rankings for every Screaming Frog user by default
would violate the package's faithful-default rule. Ask for it explicitly.

```{r, eval = FALSE}
links <- screaming_frog_links("all_inlinks.csv")
pagerank(links$edges, container_col = "container")
```

## See also

- `vignette("presets")` — region weighting and the `"content"` preset.
- `?pagerank` for the argument reference.
- `?sf_container_from_path`, `?screaming_frog_links` for the Screaming Frog path.
