---
title: "Using the pagerankr Package"
author: "Bart Turczynski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using the pagerankr Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pagerankr)
```

## Introduction

The `pagerankr` package provides a toolkit for calculating PageRank scores from web crawl data, specifically edge lists (showing links between pages) and redirect reports. This is a common task in SEO (Search Engine Optimization) analysis to understand the link equity distribution within a website.

The package offers both a main wrapper function, `pagerank()`, for a quick end-to-end calculation, and a suite of modular, pipeable functions that allow for more granular control over each step of the process. All data manipulation is done using base R, with `igraph` used for the core PageRank algorithm and `rurl` for URL canonicalization.

## Installation

If you haven't installed `pagerankr` yet, you can install it from its GitHub repository:

```r
# install.packages("devtools") # If you don't have devtools
# devtools::install_github("bart-turczynski/pagerankr")
```

Load the package:
```{r load_package}
library(pagerankr)
```

## Quick Start: Using the `pagerank()` Wrapper

The `pagerank()` function is the easiest way to get started. It orchestrates all the necessary steps: URL cleaning, redirect resolution, edge deduplication, isolate handling, and the final PageRank computation.

Let's create some sample data:

```{r sample_data_wrapper}
edges_raw <- data.frame(
  source_page = c(
    "http://example.com/home", 
    "example.com/pageA",       # No scheme
    "HTTPS://Example.com/PAGEA", # Different case, scheme
    "http://example.com/pageB?param=1#fragment",
    "http://example.com/pageC",
    "http://example.com/pageD" # pageD is a self-loop target if clean
  ),
  destination_page = c(
    "example.com/pageA",
    "http://example.com/pageB?param=1",
    "http://example.com/pageC",
    "http://example.com/home",
    "http://example.com/pageC", # Self-loop C->C
    "HTTP://EXAMPLE.COM/PAGED"  # pageD self-loop after cleaning
  ),
  stringsAsFactors = FALSE
)

redirects_report <- data.frame(
  original_url = c("http://example.com/pageB?param=1#fragment", "example.com/old-page"),
  final_url = c("http://example.com/pageB-final", "http://example.com/pageA"),
  stringsAsFactors = FALSE
)
```

Now, let's calculate PageRank using the wrapper:

```{r wrapper_example}
pr_scores <- pagerank(
  edge_list_df = edges_raw,
  redirects_df = redirects_report,
  edge_from_col = "source_page", 
  edge_to_col = "destination_page",
  redirect_from_col = "original_url",
  redirect_to_col = "final_url",
  self_loops = "drop",        # Default: drop self-loops after resolution
  drop_isolates_flag = TRUE, # Default: drop nodes with no links after processing
  damping = 0.85             # Default damping factor
)

print(pr_scores)

# Sum of PageRank should be 1 (or 0 if graph is empty after processing)
if(nrow(pr_scores) > 0) {
  cat("\nSum of PageRank scores:", sum(pr_scores$pagerank), "\n")
}
```

The `pagerank()` function automatically cleaned the URLs, resolved redirects (e.g., `pageB?param=1#fragment` should have been mapped to `pageB-final`), handled duplicate edges that might arise after cleaning/resolution, dropped self-loops, and excluded any isolated nodes before computing the PageRank.

> **Note:** As part of the workflow, any edge where either `from` or `to` is `NA` is automatically dropped before deduplication and PageRank calculation. Self-referencing redirects (where `from == to`) are also filtered out before redirect resolution.

## Modular Approach: Step-by-Step Calculation

For more control, you can use the individual functions that `pagerank()` wraps. This is useful for inspecting intermediate results or customizing parts of the workflow.

Let's use the same raw data as above.

### 1. Clean URLs

First, we clean the URLs in our edge list and redirect report. `clean_url_columns` uses `rurl::get_clean_url` internally and applies memoization for efficiency.

```{r modular_clean}
# Define parameters for rurl::get_clean_url if needed
# Default behavior of rurl::get_clean_url is often sufficient (e.g., lowercases domain, normalizes scheme)
# For this example, we'll use defaults by not passing rurl_params to pagerank(),
# or by calling clean_url_columns without extra `...` args for rurl::get_clean_url.

cleaned_edges <- clean_url_columns(
  data_frame = edges_raw, 
  columns = c("source_page", "destination_page")
)

cleaned_redirects <- clean_url_columns(
  data_frame = redirects_report, 
  columns = c("original_url", "final_url")
)

cat("Cleaned Edges:\n")
print(cleaned_edges)
cat("\nCleaned Redirects:\n")
print(cleaned_redirects)
```

*Note on Shared Memoization*: The `pagerank()` wrapper can ensure that a single URL cleaning cache is used across both the edge list and redirect list if `clean_edge_urls = TRUE` and `clean_redirect_urls = TRUE`. When calling `clean_url_columns` separately like this, each call will use its own internal memoization cache by default.

### 2. Resolve Redirects

Next, update the edge list using the cleaned redirect rules.

```{r modular_redirects}
resolved_edges <- resolve_redirects(
  edge_list_df = cleaned_edges, 
  redirects_df = cleaned_redirects,
  edge_from_col = "source_page", 
  edge_to_col = "destination_page",
  redirect_from_col = "original_url",
  redirect_to_col = "final_url"
)

cat("Edges after redirect resolution:\n")
print(resolved_edges)
```

### 3. Get Unique Edges (and Handle Self-Loops)

Remove duplicate edges and decide how to handle self-loops (e.g., pageC -> pageC).

```{r modular_unique_edges}
# Option 1: Drop self-loops (default for get_unique_edges)
unique_edges_dropped_sl <- get_unique_edges(
  edge_list_df = resolved_edges,
  self_loops = "drop",
  from_col = "source_page",
  to_col = "destination_page"
)
cat("Unique edges (self-loops dropped):\n")
print(unique_edges_dropped_sl)

# Option 2: Keep self-loops
unique_edges_kept_sl <- get_unique_edges(
  edge_list_df = resolved_edges,
  self_loops = "keep",
  from_col = "source_page",
  to_col = "destination_page"
)
cat("\nUnique edges (self-loops kept):\n")
print(unique_edges_kept_sl)
```

For the rest of this modular example, let's proceed with `unique_edges_dropped_sl`.

### 4. Handle Isolated Nodes

Decide if nodes that have no connections (isolates) after the previous steps should be included in the PageRank calculation. `drop_isolates()` returns a list of vertex names.

```{r modular_isolates}
# Option 1: Drop isolates (i.e., only keep nodes with degree > 0)
vertices_to_keep <- drop_isolates(
  edge_list_df = unique_edges_dropped_sl, 
  drop = TRUE, 
  from_col = "source_page", 
  to_col = "destination_page", 
  node_col_name = "node_name"
)
cat("Vertices for graph (isolates dropped):\n")
print(vertices_to_keep)

# Option 2: Keep all unique nodes mentioned in the (processed) edge list
# These will become graph vertices; isolates get base PageRank from teleportation.
all_vertices <- drop_isolates(
  edge_list_df = unique_edges_dropped_sl, 
  drop = FALSE, 
  from_col = "source_page", 
  to_col = "destination_page", 
  node_col_name = "node_name"
)
cat("\nAll unique vertices from processed edges (isolates included if any):\n")
print(all_vertices)
```

Let's use `vertices_to_keep` for PageRank calculation.
If `vertices_to_keep` is empty (e.g., all edges were self-loops and dropped), then `compute_pagerank` will produce an empty result.

### 5. Compute PageRank

Finally, compute PageRank scores using the processed edge list and the defined set of vertices.

```{r modular_compute_pr}
# Ensure vertices_to_keep is not NULL or empty in a way compute_pagerank expects
final_vertices_df <- NULL
if (nrow(vertices_to_keep) > 0) {
  final_vertices_df <- vertices_to_keep
}

final_pr_scores <- compute_pagerank(
  edge_list_df = unique_edges_dropped_sl, 
  vertices_df = final_vertices_df, # Can be NULL if all nodes from edges are desired
  damping = 0.85,
  from_col = "source_page",
  to_col = "destination_page",
  vertex_col_name = "node_name" # Must match column name in final_vertices_df if provided
)

cat("Final PageRank scores from modular approach:\n")
print(final_pr_scores)
if(nrow(final_pr_scores) > 0) {
  cat("\nSum of PageRank scores:", sum(final_pr_scores$pagerank), "\n")
}
```

## Using the Native Pipe `|>`

The modular functions are designed to be pipeable:

```{r modular_pipe_example, eval=FALSE}
# Note: eval=FALSE as this is illustrative of the piping syntax.
# The actual execution depends on the state of variables from previous chunks.

# Assume edges_raw and redirects_report are defined

pr_via_pipe <- edges_raw |>
  clean_url_columns(columns = c("source_page", "destination_page")) |>
  (\(df) resolve_redirects( # Lambda to handle multiple arguments for resolve_redirects
    edge_list_df = df, 
    redirects_df = clean_url_columns(redirects_report, columns = c("original_url", "final_url")),
    edge_from_col = "source_page", 
    edge_to_col = "destination_page",
    redirect_from_col = "original_url",
    redirect_to_col = "final_url"
  ))() |> # Immediately invoke the lambda
  get_unique_edges(self_loops = "drop", from_col = "source_page", to_col = "destination_page") |>
  (\(df) {
    vertices <- drop_isolates(df, drop = TRUE, from_col = "source_page", to_col = "destination_page", node_col_name = "node_name")
    compute_pagerank(
      edge_list_df = df, 
      vertices_df = if(nrow(vertices)>0) vertices else NULL, 
      from_col = "source_page", 
      to_col = "destination_page", 
      vertex_col_name = "node_name"
    )
  })()

# print(pr_via_pipe)
```
*Note*: Piping `resolve_redirects` and `compute_pagerank` requires a bit more care due to their multiple primary data frame arguments (`edge_list_df`, `redirects_df` for the former; `edge_list_df`, `vertices_df` for the latter). Lambdas or intermediate assignments might be needed for clarity, as shown in the `pagerank()` wrapper itself.

## Key Parameters and Choices

*   **URL Cleaning (`clean_url_columns`, `pagerank(rurl_params=...)`)**: 
    *   Uses `rurl::get_clean_url`. You can pass parameters to `rurl::get_clean_url` either via `...` in `clean_url_columns` or via the `rurl_params` list argument in `pagerank()`.
    *   The `pagerank()` wrapper handles whether to clean edge list URLs (`clean_edge_urls`) and redirect URLs (`clean_redirect_urls`) and can use a shared cache.
*   **Self-Loops (`get_unique_edges`, `pagerank(self_loops=...)`)**: Choose to "drop" (default) or "keep" self-loops.
*   **Isolates (`drop_isolates`, `pagerank(drop_isolates_flag=...)`)**: Decide whether to include nodes with no connections in the final PageRank graph. If kept, they receive a base PageRank score due to the damping factor.
*   **Damping Factor (`compute_pagerank`, `pagerank(damping=...)`)**: The `damping` parameter (typically 0.85) for the PageRank algorithm can be adjusted.

This vignette provides an overview of how to use the `pagerankr` package. For more detailed information on each function and its arguments, please consult the package documentation (e.g., `?pagerank`, `?clean_url_columns`). 

> **Note:** Edges with `NA` in either endpoint are dropped automatically by the workflow. Self-referencing redirects are filtered out before redirect resolution. 