---
title: "Using the pagerankr Package"
author: "Bart Turczynski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using the pagerankr Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pagerankr)
```

## Introduction

The `pagerankr` package provides an SEO-focused toolkit for calculating,
comparing, and analysing PageRank scores from web crawl data. It handles
the full pipeline from raw crawl exports (edge lists, redirect reports,
nofollow annotations, indexability status) to actionable insights.

The package offers:

- **`pagerank()`** -- end-to-end wrapper for quick calculations
- **Modular functions** -- for granular control over each pipeline step
- **Model comparison** -- grid search, diff, and distribution metrics
- **What-if simulation** -- test link/redirect changes before production

All data manipulation uses base R. `igraph` powers the PageRank algorithm
and graph-based redirect resolution. `rurl` handles URL canonicalisation.

## Installation

```r
# install.packages("devtools")
devtools::install_github("bart-turczynski/pagerankr")
```

```{r load_package}
library(pagerankr)
```

## Quick Start

```{r quick_start}
edges <- data.frame(
  from = c(
    "http://example.com/home", "http://example.com/about",
    "http://example.com/blog", "http://example.com/blog"
  ),
  to = c(
    "http://example.com/about", "http://example.com/home",
    "http://example.com/home", "http://example.com/about"
  ),
  stringsAsFactors = FALSE
)

pr <- pagerank(edges)
print(pr)
```

## Redirect Resolution

### Basic redirect handling

```{r redirects_basic}
edges <- data.frame(
  from = c("A", "B", "C"),
  to = c("B", "C", "D"),
  stringsAsFactors = FALSE
)
redirects <- data.frame(
  from = c("B", "C"),
  to = c("B_new", "C_new"),
  stringsAsFactors = FALSE
)

# resolve_redirects() applies redirect rules to an edge list
resolved <- resolve_redirects(edges, redirects)
print(resolved)
```

Redirect chains are resolved transitively: if A redirects to B and B
redirects to C, then A resolves to C.

### Conflicting redirects

Crawl data often has the same URL redirecting to different targets.
Use `duplicate_from_policy` to control the behaviour:

```{r redirects_conflict}
redirects_conflict <- data.frame(
  from = c("old", "old", "old"),
  to = c("target_A", "target_B", "target_B"),
  stringsAsFactors = FALSE
)

# "most_frequent" picks the most common target
edges_simple <- data.frame(from = "X", to = "old", stringsAsFactors = FALSE)
resolve_redirects(edges_simple, redirects_conflict,
  duplicate_from_policy = "most_frequent"
)
```

Available policies: `strict` (default, errors on conflict), `first_wins`,
`last_wins`, `most_frequent`, `prune_source`, `resolve_if_consistent`.

### Redirect loops

Redirect cycles (A -> B -> C -> A) are detected via strongly connected
components. Use `loop_handling` to control behaviour:

```{r redirects_loops}
edges_loop <- data.frame(from = "X", to = "A", stringsAsFactors = FALSE)
redirects_loop <- data.frame(
  from = c("A", "B", "C", "D"),
  to = c("B", "C", "A", "E"),
  stringsAsFactors = FALSE
)

# "prune_loop" removes cycle edges; linear redirects still work
resolved <- resolve_redirects(edges_loop, redirects_loop,
  loop_handling = "prune_loop"
)
print(resolved)
```

Available policies: `error` (default), `prune_loop`, `break_arrow`
(keeps the highest in-degree node as sink).

### Standalone link resolution

Use `resolve_links()` to apply redirects and deduplicate without
computing PageRank -- useful for inspecting the resolved graph:

```{r resolve_links_example}
edges <- data.frame(
  from = c("A", "A", "B"),
  to = c("B", "B", "C"),
  stringsAsFactors = FALSE
)
redirects <- data.frame(from = "B", to = "B_final", stringsAsFactors = FALSE)

resolve_links(edges, redirects, clean_urls = FALSE)
```

## Nofollow Handling

`pagerankr` provides three explicit propagation policies for links marked
nofollow:

- `"evaporate"` (the default) keeps each nofollow link as an outgoing slot.
  It consumes its weighted share of the source page's budget, but that share
  reaches a sink rather than the target.
- `"drop"` removes nofollow links before allocating the outgoing budget. They
  consume no slots, and followed links divide the full budget.
- `"keep"` follows the links normally, so their targets receive their shares.

These are package modeling choices; they are not claims about a search
engine's current implementation.

```{r nofollow}
edges <- data.frame(
  from = c("Hub", "Hub", "Hub"),
  to = c("A", "B", "C"),
  nofollow = c(FALSE, FALSE, TRUE),
  stringsAsFactors = FALSE
)

# "evaporate": PR splits 3 ways, C's share vanishes
pr_evap <- pagerank(edges,
  nofollow_col = "nofollow",
  nofollow_action = "evaporate",
  clean_edge_urls = FALSE
)
print(pr_evap)

# "drop": nofollow edges removed, PR splits only among followed links
pr_drop <- pagerank(edges,
  nofollow_col = "nofollow",
  nofollow_action = "drop",
  clean_edge_urls = FALSE
)
print(pr_drop)
```

## Indexability

### noindex pages

`pagerankr` assumes that the ranked corpus consists of indexed documents.
Under this indexed-corpus assumption, a `noindex` page is outside the ranked
corpus. It may receive authority through inlinks, but cannot redistribute that
authority within the indexed graph. The package therefore treats its outgoing
links as nofollow for propagation, applying the selected `nofollow_action`:
slot-consuming `"evaporate"`, slot-removing `"drop"`, or normally followed
`"keep"`.

This is strictly a `pagerankr` PageRank assumption. It does not mean that
Google or another search engine defines `noindex` as an explicit nofollow
directive. A noindex page can remain visible in the result to make its received
authority auditable; hiding it would be an optional reporting choice, not a
change required for correct propagation under this model.

```{r indexability_noindex}
edges <- data.frame(
  from = c("Home", "NoindexPage", "NoindexPage"),
  to = c("NoindexPage", "A", "B"),
  stringsAsFactors = FALSE
)

idx_status <- data.frame(
  url = "NoindexPage",
  indexability_status = "noindex",
  stringsAsFactors = FALSE
)

pr <- pagerank(edges,
  indexability_df = idx_status,
  clean_edge_urls = FALSE
)
print(pr)
```

### robots.txt blocked pages

A robots.txt-blocked page receives inbound PageRank but cannot pass it
outbound (Google can't see the links). The PR is either trapped or vanishes:

```{r indexability_robots}
edges <- data.frame(
  from = c("Home", "Blocked", "A"),
  to = c("Blocked", "A", "Home"),
  stringsAsFactors = FALSE
)

idx_status <- data.frame(
  url = "Blocked",
  indexability_status = "Blocked by robots.txt",
  stringsAsFactors = FALSE
)

# "trap": blocked page keeps its PR (visible in results)
pr_trap <- pagerank(edges,
  indexability_df = idx_status,
  robots_blocked_action = "trap",
  clean_edge_urls = FALSE
)
print(pr_trap)

# "vanish": blocked page removed from results
pr_vanish <- pagerank(edges,
  indexability_df = idx_status,
  robots_blocked_action = "vanish",
  clean_edge_urls = FALSE
)
print(pr_vanish)
```

## Weighted Edges

Pass a `weight_col` to use link weights (e.g., number of links, link
position scores) in the PageRank calculation:

```{r weighted}
edges <- data.frame(
  from = c("A", "A", "B"),
  to = c("B", "C", "C"),
  weight = c(3, 1, 1),
  stringsAsFactors = FALSE
)

pr <- pagerank(edges, weight_col = "weight", clean_edge_urls = FALSE)
print(pr)
```

## Domain Filtering

`filter_links_by_domain()` scopes edge lists by domain, useful for
separating internal vs. external links:
```{r domain_filter}
edges <- data.frame(
  from = c(
    "http://example.com/a", "http://example.com/b",
    "http://other.com/c"
  ),
  to = c(
    "http://example.com/b", "http://other.com/d",
    "http://example.com/a"
  ),
  stringsAsFactors = FALSE
)

# Keep only internal links
internal <- filter_links_by_domain(edges, keep_domains = "example.com")
print(internal)
```

## HITS hub and authority

`hits()` computes Kleinberg's HITS hub and authority scores over the *same*
cleaned, redirect/canonical-folded, domain-filtered, deduplicated link graph as
`pagerank()`. Because both share the identity pipeline, you can join hub,
authority, and PageRank on `node_name` directly.

The two scores reinforce each other: a good **authority** is pointed to by good
**hubs**, and a good hub points to good authorities. Formally, with adjacency
matrix `A`, authority is the dominant eigenvector of `A^T A` and hub is the
dominant eigenvector of `A A^T`.

```{r hits}
edges <- data.frame(
  from = c("A", "A", "B"),
  to = c("B", "C", "C"),
  stringsAsFactors = FALSE
)

# A only points out (pure hub); C is only pointed to (pure authority).
hits(edges, clean_edge_urls = FALSE)
```

Scores are scaled so each maximum is `1` (`scale = TRUE`, the conventional HITS
reporting). HITS already captures both directions of authority flow, so there
is no `reverse` flag, and the PageRank-only forward-flow devices (nofollow
evaporation, indexability transforms, the TIPR teleport prior) are not exposed.

**Whole-graph caveat.** Kleinberg's original HITS ran on a small, query-focused
base set of pages. `hits()` runs on the full (or domain-filtered) site graph
`pagerankr` assembles, so treat the scores as site-wide structural
centralities rather than query-relevance scores.

## Model Comparison

### Comparing two models

```{r compare}
edges <- data.frame(
  from = c("A", "B", "C", "D"),
  to = c("B", "C", "D", "A"),
  stringsAsFactors = FALSE
)

pr_85 <- pagerank(edges, damping = 0.85, clean_edge_urls = FALSE)
pr_90 <- pagerank(edges, damping = 0.90, clean_edge_urls = FALSE)

diff <- compare_pagerank(pr_85, pr_90, label_a = "d=0.85", label_b = "d=0.90")
print(diff)
cat("\nCorrelation summary:\n")
print(attr(diff, "summary"))
```

### Parameter grid search

Run PageRank across multiple parameter combinations to find the most
informative model:

```{r grid}
edges <- data.frame(
  from = c("A", "A", "B", "C"),
  to = c("B", "C", "C", "A"),
  stringsAsFactors = FALSE
)

grid <- auto_grid(damping = c(0.75, 0.85, 0.95))
results <- pagerank_grid(edges, params_grid = grid, clean_edge_urls = FALSE)

# Analyse distribution metrics across the grid
analysis <- analyze_pagerank_grid(results)
print(analysis)
```

### Distribution metrics

Standalone metrics for any PageRank vector:

```{r metrics}
pr <- pagerank(edges, clean_edge_urls = FALSE)
cat("Gini coefficient:", pr_gini(pr$pagerank), "\n")
cat("Entropy:", pr_entropy(pr$pagerank), "\n")
cat("Top-1 share:", pr_top_k_share(pr$pagerank, k = 1), "\n")
```

## Simulating Changes

A common SEO workflow is asking "what happens to PageRank if I add these
links, remove those links, or implement these redirects?"

### Adding links

```{r simulate_add_link}
site_links <- data.frame(
  from = c("Home", "Home", "About", "Blog"),
  to = c("About", "Blog", "Home", "Home"),
  stringsAsFactors = FALSE
)

impact <- simulate_changes(
  site_links,
  add_links_df = data.frame(
    from = "Blog", to = "About",
    stringsAsFactors = FALSE
  ),
  clean_edge_urls = FALSE
)
print(impact)
```

### Removing links

```{r simulate_remove_link}
impact_remove <- simulate_changes(
  site_links,
  remove_links_df = data.frame(
    from = "Home", to = "Blog",
    stringsAsFactors = FALSE
  ),
  clean_edge_urls = FALSE
)
print(impact_remove)
```

### Simulating redirects

```{r simulate_redirect}
extended_links <- rbind(site_links, data.frame(
  from = "OldPage", to = "Home", stringsAsFactors = FALSE
))

impact_redirect <- simulate_changes(
  extended_links,
  add_redirects_df = data.frame(
    from = "OldPage", to = "About",
    stringsAsFactors = FALSE
  ),
  clean_edge_urls = FALSE
)
print(impact_redirect)
```

### Combined changes

```{r simulate_combined}
impact_all <- simulate_changes(
  site_links,
  add_links_df = data.frame(
    from = "Blog", to = "About",
    stringsAsFactors = FALSE
  ),
  remove_links_df = data.frame(
    from = "Home", to = "Blog",
    stringsAsFactors = FALSE
  ),
  clean_edge_urls = FALSE
)
print(impact_all)
```

All `pagerank()` parameters (damping, nofollow, indexability, etc.) can
be passed through `simulate_changes()` via `...`.

## Function Reference

| Function | Purpose |
|---|---|
| `pagerank()` | End-to-end: clean, resolve, compute |
| `hits()` | End-to-end HITS hub + authority on the same graph |
| `compute_hits()` | Low-level igraph HITS wrapper |
| `resolve_links()` | Resolve redirects + deduplicate (no PR) |
| `simulate_changes()` | What-if: compare baseline vs. proposed |
| `compare_pagerank()` | Diff two PR results with rank shifts |
| `pagerank_grid()` | Run PR across parameter combinations |
| `auto_grid()` | Generate parameter grid |
| `analyze_pagerank_grid()` | Distribution metrics across a grid |
| `filter_links_by_domain()` | Scope edges by domain |
| `resolve_redirects()` | Apply redirects to an edge list |
| `clean_url_columns()` | Canonicalise URLs |
| `get_unique_edges()` | Deduplicate, handle self-loops |
| `drop_isolates()` | Include/exclude disconnected nodes |
| `compute_pagerank()` | Low-level igraph PR wrapper |
| `pr_gini()` | Gini coefficient of PR distribution |
| `pr_entropy()` | Entropy of PR distribution |
| `pr_top_k_share()` | Top-k concentration of PR |
