---
title: "Presets: named views of the link graph"
author: "Bart Turczynski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Presets: named views of the link graph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pagerankr)
```

## Why presets exist

PageRank is a lens, not the deliverable. The same crawl answers different
questions depending on how you prepare the graph first: *what does the site
look like exactly as crawled?* *What does it look like once its own canonical
and nofollow declarations are honored?* *Which pages feed authority outward
rather than collect it?* *What survives once site chrome stops voting?*

`pagerank()` can express all of those, but only through a long argument list.
A **preset** is a small named bundle of those arguments describing one
recurring view, so a view becomes a one-liner:

```{r}
edges <- data.frame(
  from = c("A", "A", "B", "C"),
  to = c("B", "C", "A", "A")
)

pagerank(edges, preset = "raw")
```

Presets are opt-in. `pagerank()` with no `preset` behaves exactly as it always
has — nothing here changes a default.

## The four presets

```{r}
pr_preset("raw")
```

A preset is a plain named list, so you can always print one to see exactly what
it does. Nothing is hidden: the table below is the whole registry.

| Preset | Expands to | The view |
|---|---|---|
| `"raw"` | `self_loops = "keep"`, `drop_isolates_flag = FALSE`, `nofollow_action = "keep"`, `out_of_scope_fold = "keep"` | The graph exactly as crawled. Nothing is applied; `rel=nofollow` votes like any other link. The faithful baseline. |
| `"declared"` | `self_loops = "drop"`, `drop_isolates_flag = TRUE`, `nofollow_action = "evaporate"`, `robots_blocked_action = "show"`, `out_of_scope_fold = "relabel"` | The graph after honoring what the site *declares* — nofollow, canonicals, redirects, robots. A **pure pin of today's defaults**. |
| `"reversed"` | `reverse = TRUE` | Every edge flipped: a page scores highly when it points *at* well-connected pages. The feeder view. |
| `"content"` | `placement_weights = c(content = 1, nav = 0.1, header = 0.1, footer = 0.1, aside = 0.1)` | Links in the main content keep their full vote; links in site chrome are discounted to a tenth. |

### `"raw"` — the faithful baseline

The naive view, and deliberately the one the package makes easiest to state.
Every other view is a claim about what *should* count; `"raw"` is the claim-free
comparison point. Reach for it when you want to see how much of a ranking your
own preparation choices produced.

### `"declared"` — a pin, not a change

`"declared"` sets every value to what `pagerank()` already defaults to. Running
with it produces byte-identical scores to running without it:

```{r}
identical(
  pagerank(edges, preset = "declared")$pagerank,
  pagerank(edges)$pagerank
)
```

That is the point. Its value is *provenance*: a run made with
`preset = "declared"` is a recorded, auditable statement that the declared view
was intended, and it stays pinned to the bundle as documented even if a future
package default moves.

Note the boundary: a preset sets **policy, never data**. `"declared"` turns on
*honoring* canonicals and redirects, but the canonical and redirect tables
themselves (`canonicals_df`, `redirects_df`, `indexability_df`) are still yours
to supply.

### `"reversed"` — the feeder view

```{r}
pagerank(edges, preset = "reversed")
```

`topic_feeder_pagerank()` already reverses the graph itself, so
`preset = "reversed"` is a harmless no-op there rather than an error.

### `"content"` — stop letting the navigation rank the site

Site chrome — navigation, header, footer, sidebar — is typically the large
majority of a crawl's edges. Left unweighted it does not merely influence the
ranking, it *manufactures* it: every page links to the same handful of chrome
targets, so those targets win by sheer repetition regardless of editorial
intent.

`"content"` discounts them. It needs one thing from you: `placement_col`, the
name of a column saying which page region each link was found in. That is data,
so the preset cannot supply it.

```{r}
placed <- data.frame(
  from = c("A", "A", "B", "C"),
  to = c("B", "C", "A", "A"),
  region = c("content", "footer", "content", "content")
)

pagerank(placed, preset = "content", placement_col = "region")
```

`B` and `C` are each linked once from `A`. Unweighted they tie; weighted, the
footer link is worth a tenth of the in-content one and `C` falls behind.

The accepted region terms are `content`, `nav`, `header`, `footer` and `aside`.
`pagerank_screaming_frog()` derives them for you from the crawl, so there
`preset = "content"` works with no extra argument.

Two properties worth stating explicitly:

- **All five terms are named.** Regions absent from `placement_weights` keep
  weight 1, which is what makes a partial recipe like `c(nav = 0.1)` mean
  "discount nav, leave everything else alone". A preset is a *complete* recipe,
  so it names every term — otherwise footer and aside would quietly outweigh
  nav tenfold.
- **Edges are downweighted, never dropped.** Region detection is a heuristic,
  so a misclassified content link at 0.1 is a small error where a dropped one
  is a silent deletion. Dropping most of a graph also manufactures isolates and
  dangling pages, at which point you are measuring your own sink policy rather
  than the site.

## Precedence: explicit argument > preset > default

A preset never silently overrides something you typed.

```{r}
# The raw bundle says nofollow_action = "keep"; the explicit argument wins.
attr(
  pagerank(edges, preset = "raw", nofollow_action = "drop"),
  "transition_audit"
)$config$nofollow_action
```

This holds through every wrapper that forwards `...` to `pagerank()` —
`trustrank()`, `topic_sensitive_pagerank()`, `topic_feeder_pagerank()` and
`pagerank_screaming_frog()` — with one boundary: arguments a wrapper sets
*itself* are wrapper-owned, and a preset cannot change those.

Presets are not composable with each other; `preset` takes a single bundle.
They do not need to be: `"content"` sets only placement weights and leaves
every hygiene knob at its default, and the defaults *are* the `"declared"`
view.

## Inspecting and adapting a bundle

Because a preset is an ordinary list, you can splice it, print it, or use it as
a starting point:

```{r}
# Splice a bundle into a call
do.call(pagerank, c(list(edges), pr_preset("raw")))

# Start from a preset and adjust
my_view <- pr_preset("raw")
my_view$drop_isolates_flag <- TRUE
pagerank(edges, preset = my_view)
```

Hand-rolled bundles are validated against `pagerank()`'s formals, so a
misspelled argument fails loudly instead of sliding through `...`.

## Provenance in the transition audit

The audit attached to every result records which named view produced it:

```{r}
attr(pagerank(edges, preset = "raw"), "transition_audit")$config$preset
```

It records the preset name for a registered preset (whether passed by name or
as a `pr_preset()` result), `"custom"` for a hand-rolled bundle, and `NULL`
when no preset was used. Two runs that expand to the same configuration are
therefore still distinguishable by the intent behind them — which, for a pure
pin like `"declared"`, is the only durable record that intent existed.

## See also

- `?pr_preset` for the reference documentation.
- `vignette("pagerankr-usage")` for the underlying arguments each preset sets.
