---
title: "TrustRank: seed-biased PageRank in pagerankr"
author: "Bart Turczynski"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{TrustRank: seed-biased PageRank in pagerankr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(pagerankr)
```

## What TrustRank is

**TrustRank** (Gyöngyi, Garcia-Molina & Pedersen, 2004) is ordinary PageRank
with one change: instead of the random surfer teleporting to a *uniform* random
page, it teleports back into a set of **trusted seed** pages. Trust then
propagates outward along links and *attenuates with distance* — the further a
page is from the trusted core, the less trust reaches it. The PageRank damping
factor is exactly that attenuation mechanism, so no extra machinery is needed.

The original motivation was web spam: pick a small set of human-vetted "good"
pages, and let trust flow from them so that link farms — which good pages rarely
link to — stay low. The same mechanic is useful on a single site for any
"authority radiates from these known-good pages" question: your editorial hubs,
your hand-picked cornerstone content, or a manually curated quality set.

This is **seed-biased PageRank**, not a spam classifier. `pagerankr` supplies
the biased-propagation core; *which* pages are trustworthy is your call.

## It reuses the personalization path — no new solver

A trusted-seed set is just a teleport (personalization) prior. `pagerankr`
already builds personalized PageRank from a per-URL prior via
`pagerank(prior_df = ...)` (the TIPR path). TrustRank is therefore two small
helpers over that engine:

- `trust_seed_prior()` turns a seed set into a `prior_df`.
- `trustrank()` is the one-call convenience wrapper.

## A worked example

```{r}
edges <- data.frame(
  from = c("/", "/", "/hub", "/hub", "/good", "/spam", "/spam"),
  to = c("/hub", "/good", "/good", "/deep", "/hub", "/sink", "/good"),
  stringsAsFactors = FALSE
)
```

Here `/` and `/hub` are our trusted editorial core. `/spam` links into the good
neighborhood (a classic spam tactic) but nothing trusted links *to* it, and it
feeds an off-topic `/sink`.

### Build the seed prior explicitly

```{r}
prior <- trust_seed_prior(c("/", "/hub"))
prior
```

Equal weights reproduce TrustRank's uniform distribution over the trusted set.
Run it through `pagerank()` like any other prior:

```{r}
pr_manual <- pagerank(edges, prior_df = prior, clean_edge_urls = FALSE)
pr_manual[order(-pr_manual$pagerank), c("node_name", "pagerank")]
```

### Or use the convenience wrapper

```{r}
tr <- trustrank(edges, c("/", "/hub"), clean_edge_urls = FALSE)
tr[order(-tr$pagerank), c("node_name", "pagerank", "prior_weight")]
```

`trustrank()` is identical to the manual call — it just builds the prior for
you. Two things to read in the output:

- **`prior_weight`** is the teleport mass each page receives. The seeds carry
  it; `/spam` and `/sink`, outside the trusted neighborhood, get exactly `0`.
- A non-seed page reachable *from* the trusted core (e.g. `/deep`, linked from
  `/hub`) still earns trust through inheritance, while the seed-unreachable
  `/spam` region stays suppressed relative to plain PageRank.

### Compare against uniform PageRank

```{r}
uni <- pagerank(edges, clean_edge_urls = FALSE)
compare_pagerank(uni, tr)[, c("node_name", "pagerank_a", "pagerank_b", "delta")]
```

`pagerank_a` is uniform PageRank, `pagerank_b` is TrustRank. The trusted core
and what it links to gain; the untrusted region loses.

## Graded trust and a teleport floor

Trust need not be all-or-nothing. Give some seeds more weight than others:

```{r}
trustrank(
  edges,
  data.frame(url = c("/", "/hub"), weight = c(3, 1)),
  clean_edge_urls = FALSE
)[, c("node_name", "pagerank", "prior_weight")]
```

By default untrusted, seed-unreachable pages receive no teleport mass at all.
To give every page a small floor (a blend of trust teleport and uniform
teleport), pass `prior_alpha`:

```{r}
trustrank(
  edges, c("/", "/hub"),
  prior_alpha = 0.15, clean_edge_urls = FALSE
)[, c("node_name", "pagerank", "prior_weight")]
```

`prior_alpha = 0` (the default) is pure trust teleport; `prior_alpha = 1`
reproduces uniform PageRank.

## Notes

- Seed weights are an **additive trust budget**: if two seed URLs fold onto the
  same vertex (redirect or canonical variants), their weights sum — consistent
  with the prior contract in `pagerank()` and `align_prior_to_vertices()`.
- Everything `pagerank()` accepts flows through `trustrank()` via `...`:
  redirects, canonicals, URL cleaning, domain/host filtering, edge weights, and
  duplicate-edge policy. Trusted seeds are canonicalized and folded into the
  same vertex namespace as the edges before alignment.
- For *topic*-biased authority (multiple content clusters, blended), see
  `topic_sensitive_pagerank()`; it uses the same personalization path with a
  per-topic prior instead of a single trusted set.
